Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
cite cleaned up a bit

Recent work by~\citet{miltner+:bijective-synthesis} has also demonstrated
how techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize the simplest class of lenses: so-called {\em
  bijective lenses}, where no information is discarded when transforming
data in either direction.

@@

Initially, we hypothesized that extending the synthesis algorithms
of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
as opposed to bijective lenses 
would be relatively straightforward:  Simply replace one set of combinators with another
and execute a search with similar heuristics.  However, in our early,
naive attempts to do so, we ran
into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
The goal was to set up the idea that there are ``two parts'' to a symmetric
lens transformation: the information from $s$ and the information from $t$.  In
a future paragraph in the intro we are going to explain that the central heuristic
involves minimizing the information from $t$.}

@@

Now that we have a way to calculate likelihood, we need a way to search for
likely lenses. Previous work on synthesis~\cite{?,?,?} has shown that we can use
types to dramatically restrict the search space and improve the effectiveness of
synthesis. In our setting, types are pairs of stochastic regular expressions,
which come with a rich algebraic structure that makes each regular expression
semantically equal to infinitely many others. To tame that search space and
leverage the benefits of type-directed synthesis, we split our algorithm into
two communicating synthesizers following earlier
work~\cite{quotient-lens-synthesis}. The first synthesizer, \Expand, navigates
the space of semantically equivalent regular expressions by applying rewrite
rules that preserve both the semantics and the probability distributions of
those stochastic regular expressions. Our algorithm ranks pairs of stochastic
regular expressions by the number of rewrite rule applications required to
obtain the pair from the stochastic regular expressions given as input. It
passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

@@

% begin related-work
\section{Related Work}

@@

\subsection{Data Transformation Synthesis}
Over the past decade, the programming languages community has explored
the synthesis of programs from a wide variety of angles.  One of
the key ideas is typically to narrow the program search space by focusing
on a specific domain, and imposing constraints on syntax~\cite{sygus} or
typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,}, or both.
%, and as has been advocated
%more generally under the rubric of syntax-guided synthesis~\cite{sygus}.

Automation of string transformations, in particular, has been the focus
of much prior attention.  For example, Gulwani and others work on FlashFill
generates one-way spreadsheet transformations from input/output
examples~\cite{flashfill,flashextract}.  On the one hand,
FlashFill is easier to use because one need not specify the type of
the data being transformed.  On the other hand, this type information
makes it possible to transform more complex formats---Flash Fill
does not synthesize programs with nested loops, for instance,
and hence is incapable of synthesizing the majority of our
benchmarks, even in one direction~\cite{flashfill}.
%% Flashfill, designed
%% for transformation of the cells of spreadsheets, does not scale to
%% more complex formats
%% While this provides an easier
%% interaction model with the user than this work, it requires the algorithm to
%% learn the data formats from the examples. This makes the task much more
%% difficult for the synthesis algorithm, particularly in addressing iterations and
%% disjunctions. Prior work has shown that learning from examples and data format
%% descriptions enables synthesis of much more complex transformations than are
%% permitted by examples alone. Recent work has gone into making these algorithms
%% more robust by delaying the selection of program until seeing what further
%% inputs the program is run on~\cite{?}. By doing this, the synthesis algorithm
%% can learn additional information about the formats from the use of the program.

>>>>

Recent work by~\citet{optician} has also demonstrated
how techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize the simplest class of lenses: so-called {\em
  bijective lenses}, where no information is discarded when transforming
data in either direction.

@@

Initially, we hypothesized that extending the synthesis algorithms
of Miltner \ETAL~\cite{optician} to operate over symmetric lenses
as opposed to bijective lenses 
would be relatively straightforward:  Simply replace one set of combinators with another
and execute a search with similar heuristics.  However, in our early,
naive attempts to do so, we ran
into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
The goal was to set up the idea that there are ``two parts'' to a symmetric
lens transformation: the information from $s$ and the information from $t$.  In
a future paragraph in the intro we are going to explain that the central heuristic
involves minimizing the information from $t$.}

@@

Now that we have a way to calculate likelihood, we need a way to search for
likely lenses. Previous work on synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16} has shown that we can use
types to dramatically restrict the search space and improve the effectiveness of
synthesis. In our setting, types are pairs of stochastic regular expressions,
which come with a rich algebraic structure that makes each regular expression
semantically equal to infinitely many others. To tame that search space and
leverage the benefits of type-directed synthesis, we split our algorithm into
two communicating synthesizers following earlier
work~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand, navigates
the space of semantically equivalent regular expressions by applying rewrite
rules that preserve both the semantics and the probability distributions of
those stochastic regular expressions. Our algorithm ranks pairs of stochastic
regular expressions by the number of rewrite rule applications required to
obtain the pair from the stochastic regular expressions given as input. It
passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

@@

% begin related-work
\section{Related Work}
\label{sec:related}

@@

\subsection{Data Transformation Synthesis}
Over the past decade, the programming languages community has explored
the synthesis of programs from a wide variety of angles.  One of
the key ideas is typically to narrow the program search space by focusing
on a specific domain, and imposing constraints on syntax~\cite{sygus} or
typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16}, or both.
%, and as has been advocated
%more generally under the rubric of syntax-guided synthesis~\cite{sygus}.

Automation of string transformations, in particular, has been the focus
of much prior attention.  For example, Gulwani and others work on FlashFill
generates one-way spreadsheet transformations from input/output
examples~\cite{flashfill,le-pldi-2014}.  On the one hand,
FlashFill is easier to use because one need not specify the type of
the data being transformed.  On the other hand, this type information
makes it possible to transform more complex formats---Flash Fill
does not synthesize programs with nested loops, for instance,
and hence is incapable of synthesizing the majority of our
benchmarks, even in one direction~\cite{flashfill}.
%% Flashfill, designed
%% for transformation of the cells of spreadsheets, does not scale to
%% more complex formats
%% While this provides an easier
%% interaction model with the user than this work, it requires the algorithm to
%% learn the data formats from the examples. This makes the task much more
%% difficult for the synthesis algorithm, particularly in addressing iterations and
%% disjunctions. Prior work has shown that learning from examples and data format
%% descriptions enables synthesis of much more complex transformations than are
%% permitted by examples alone. Recent work has gone into making these algorithms
%% more robust by delaying the selection of program until seeing what further
%% inputs the program is run on~\cite{?}. By doing this, the synthesis algorithm
%% can learn additional information about the formats from the use of the program.
